Towards understanding the central limit theorem by learning Python basics

### 2022-05-17 paper/article.tex
Simplify zip

As another activity, a third program was given (see Fig.~\ref{6c.py}) to prepare the last
topic, namely, some steps towards the central limit theorem (CLT, \cite{Polya1920}).
\begin{figure}
\lstinputlisting[language=Python]{../progs/en_US/6c.py}
\caption{A third program in the sixth sheet (translated into English)}
\label{6c.py}
\end{figure}
Here we provide the questions to this third program:
\begin{enumerate}
\item What computation is performed by this program?
\item Consider following modifications:
\begin{lstlisting}[language=Python,firstnumber=2]
Ω = set(itertools.product({1, 2, 3, 4, 5, 6}, repeat=4))
possible_values = range(30)
\end{lstlisting}
\begin{lstlisting}[language=Python,firstnumber=8]
  heads[sum(ω)] += 1
\end{lstlisting}
Explain the output of the modified program.
\item We are about to plot the obtained data. So we change line 10 as follows:
\begin{lstlisting}[language=Python,firstnumber=10]
for w in zip(*(possible_values, probabilities)):
  print(f"{w[0]},{w[1]}")
\end{lstlisting}
Copy the output of this program in a spreadsheet software and create an XY diagram
to get some graphical output.
\item Extend the program with these two additional lines:
\begin{lstlisting}[language=Python,firstnumber=12]
distribution = [sum(probabilities[:h]) for h in possible_values]
print(list(zip(*(possible_values, distribution))))
\end{lstlisting}
Explain the result.
\end{enumerate}
In fact, the variable naming was somewhat incorrect here (because \texttt{heads} 
actually refers to the sums of the outcomes), but we wanted to make very little
changes on the code.

Among the authors there was a quite a long discussion on getting the output in
the best fashion. The Python construction \texttt{list(zip(*(...)))} is a kind
of compromise between beautiful output with much code and simple output with less code,
however, this is something non-trivial in Python.
Finally a graphical output was also expected---students should be familiar with
popular spreadsheet software and visualize data in an easier way than programming.

>>>>

As another activity, a third program was given (see Fig.~\ref{6c.py}) to prepare the last
topic, namely, some steps towards the central limit theorem (CLT, \cite{Polya1920}).
\begin{figure}
\lstinputlisting[language=Python]{../progs/en_US/6c.py}
\caption{A third program in the sixth sheet (translated into English)}
\label{6c.py}
\end{figure}
Here we provide the questions to this third program:
\begin{enumerate}
\item What computation is performed by this program?
\item Consider following modifications:
\begin{lstlisting}[language=Python,firstnumber=2]
Ω = set(itertools.product({1, 2, 3, 4, 5, 6}, repeat=4))
possible_values = range(30)
\end{lstlisting}
\begin{lstlisting}[language=Python,firstnumber=8]
  heads[sum(ω)] += 1
\end{lstlisting}
Explain the output of the modified program.
\item We are about to plot the obtained data. So we change line 10 as follows:
\begin{lstlisting}[language=Python,firstnumber=10]
for w in zip(possible_values, probabilities):
  print(f"{w[0]},{w[1]}")
\end{lstlisting}
Copy the output of this program in a spreadsheet software and create an XY diagram
to get some graphical output.
\item Extend the program with these two additional lines:
\begin{lstlisting}[language=Python,firstnumber=12]
distribution = [sum(probabilities[:h]) for h in possible_values]
print(list(zip(possible_values, distribution)))
\end{lstlisting}
Explain the result.
\end{enumerate}
In fact, the variable naming was somewhat incorrect here (because \texttt{heads} 
actually refers to the sums of the outcomes), but we wanted to make very little
changes on the code.

Among the authors there was a quite a long discussion on getting the output in
the best fashion. The Python construction \texttt{list(zip(...))} is a kind
of compromise between beautiful output with much code and simple output with less code,
however, this is something non-trivial in Python.
Finally a graphical output was also expected---students should be familiar with
popular spreadsheet software and visualize data in an easier way than programming.
